---
title: "P8105 Homework 6"
author: "Hyun Kim (hk3373)"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)

theme_set(theme_minimal() + 
  theme(legend.position = "bottom", 
      plot.title = element_text(hjust = 0.5)))

knitr::opts_chunk$set(
	fig.width = 8, 
  fig.height = 6
)
```

# Problem 1

## Download 2017 Central Park weather data
```{r download_weather_data}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Use 5000 bootstrap samples and produce estimates for each boostrap sample
```{r produce_sample_estimates}
boot_strap_df =
  weather_df |>
  bootstrap(n = 5000) |>
  mutate(models = map(strap, \(df) lm(tmax ~ tmin, data = df)))

r_squared_df =
  boot_strap_df |>
  mutate(r_squared = map(models, broom::glance)) |> 
  select(-strap, -models) |> 
  unnest(r_squared)

log_beta_df = 
  boot_strap_df |>
  mutate(beta = map(models, broom::tidy)) |> 
  select(-strap, -models) |> 
  unnest(beta) |>
  select(.id, term, estimate) |>
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |>
  rename(intercept = "(Intercept)") |>
  mutate(log_beta = log(intercept * tmin))
```

## Plot distribution of the estimates

### Estimate of r^2
```{r plot_r_squared}
r_squared_df |>
  ggplot(aes(x = r.squared)) +
  geom_density() + 
  labs(y = "count",
       x = "Estimate of r^2",
       title = "Distribution of the Estimate of r^2")
```

The plot shows that the distribution of the estimate of r^2 is normal.

### Estimate of log(B0 * B1)
```{r plot_log_beta}
log_beta_df |>
  ggplot(aes(x = log_beta)) + 
  geom_density() + 
  labs(y = "count",
       x = "Estimate of log(B0 * B1)",
       title = "Distribution of the Estimate of log(B0 * B1)")
```

The plot shows that the distribution of the estimate of log(B0 * B1) is normal.

## Compute 95% confidence interval for the estimates

### CI of r^2 Estimate
```{r r_squared_CI}
r_squared_df |>
  summarize(
    r2_CI_lower = quantile(r.squared, 0.025),
    r2_CI_upper = quantile(r.squared, 0.975)
  ) |>
  knitr::kable(digits = 3)
```

### CI of log(B0 * B1) Estimate
```{r log_beta_CI}
log_beta_df |>
  summarize(
    log_CI_lower = quantile(log_beta, 0.025),
    log_CI_upper = quantile(log_beta, 0.975)
  ) |>
  knitr::kable(digits = 3)
```
